---
title: "Analysis module 3: Geographic differences in 16S microbiota composition"
author: "RoVI study"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  flexdashboard::flex_dashboard:
    self_contained: yes
    theme: flatly 
---

Sidebar {.sidebar} 
===================================== 
[Sample summary]    

[Taxon abundance profile]

[Alpha diversity]    

[Longitudinal abundance plots]    

[Beta diversity]    

[Differential taxa]     

[Longitudinal abundance comparisons]    

[Session info]





Sample summary {.hidden}
=====================================

```{r, message=FALSE}
# variable to define whether to run analyses in full (TRUE) or not (FALSE)
run_rf_full = FALSE # set as TRUE to run Random Forests in full
run_nb_full = FALSE # set as TRUE to run negative binomial models in full
iter = 20 # number of cross-validation iterations to run
adonis_iter = 999

# create output folders
if ("output_module3" %in% list.files()==FALSE) {
  system("mkdir output_module3")
  system("mkdir output_module3/RF_output_stats")
  system("mkdir output_module3/RF_output_importance")
  system("mkdir output_module3/RF_output_importance_reduced")
}

# load packages and functions
source("input/packages_and_functions.R")
source("input/RF_functions.R")

# load genus/rsv-level data - full and rarefied
load("phyloseq_files/ps7_genus.RData")
load("phyloseq_files/ps7_l_t_pr_d_dd_25k.RData")
load("phyloseq_files/ps8_genus.RData")
load("phyloseq_files/ps8_l_t_pr_d_dd_25k_rarefied.RData")

# select BS1, BS2, BS3, BS5 and MS1 samples within each phyloseq object
ps7_genus = subset_samples(ps7_genus, (!is.na(dose1_shedding) | !is.na(seroconv)) & (sample_type=="BS1" | sample_type=="BS2" | sample_type=="BS3" | sample_type=="BS5" | sample_type=="MS1")) %>% prune_taxa(taxa_sums(.) > 0, .)
ps7 = subset_samples(ps7_l_t_pr_d_dd_25k, (!is.na(dose1_shedding) | !is.na(seroconv)) & (sample_type=="BS1" | sample_type=="BS2" | sample_type=="BS3" | sample_type=="BS5" | sample_type=="MS1")) %>% prune_taxa(taxa_sums(.) > 0, .)
ps8_genus = subset_samples(ps8_genus, (!is.na(dose1_shedding) | !is.na(seroconv)) & (sample_type=="BS1" | sample_type=="BS2" | sample_type=="BS3" | sample_type=="BS5" | sample_type=="MS1")) %>% prune_taxa(taxa_sums(.) > 0, .)
ps8 = subset_samples(ps8_l_t_pr_d_dd_25k_rarefied, (!is.na(dose1_shedding) | !is.na(seroconv)) & (sample_type=="BS1" | sample_type=="BS2" | sample_type=="BS3" | sample_type=="BS5" | sample_type=="MS1")) %>% prune_taxa(taxa_sums(.) > 0, .)
#all(sample_names(ps7_genus)==sample_names(ps8_genus))

# amend taxa names to be consistent across ps7 and ps8 levels
taxa_names(ps7_genus) = data.frame(as(tax_table(ps7_genus), "matrix"))$Genus
taxa_names(ps7) = data.frame(as(tax_table(ps7), "matrix"))$tax_id
taxa_names(ps8_genus) = data.frame(as(tax_table(ps8_genus), "matrix"))$Genus
taxa_names(ps8) = data.frame(as(tax_table(ps8), "matrix"))$tax_id

# select BS1, BS2, BS3, BS5 and MS1 samples within each phyloseq object
ps7_genus_infant = subset_samples(ps7_genus, (!is.na(dose1_shedding) | !is.na(seroconv)) & (sample_type=="BS1" | sample_type=="BS2" | sample_type=="BS3" | sample_type=="BS5")) %>% prune_taxa(taxa_sums(.) > 0, .)
ps7_infant = subset_samples(ps7_l_t_pr_d_dd_25k, (!is.na(dose1_shedding) | !is.na(seroconv)) & (sample_type=="BS1" | sample_type=="BS2" | sample_type=="BS3" | sample_type=="BS5")) %>% prune_taxa(taxa_sums(.) > 0, .)

# transform to relative abundances
rps7_genus = transform_sample_counts(ps7_genus, function(x) {x/sum(x)})
rps7 = transform_sample_counts(ps7, function(x) {x/sum(x)})
rps8_genus = transform_sample_counts(ps8_genus, function(x) {x/sum(x)})
rps8 = transform_sample_counts(ps8, function(x) {x/sum(x)})
```

#### Input data: infant stools (BS1, BS2, BS3, and BS5) and maternal stool (MS1) 
* n samples = `r nsamples(ps7_genus)`    
* n taxa: genera = `r ntaxa(ps7_genus)`; variants = `r sum(taxa_sums(ps7)>0)` 
* n taxa (infant): genera = `r ntaxa(ps7_genus)`; variants = `r sum(taxa_sums(ps7)>0)` 
* n taxa (rarefied): genera = `r ntaxa(ps7_genus_infant)`; variants = `r sum(taxa_sums(ps7_infant)>0)`  
* mean read count = `r round(mean(sample_sums(ps7_genus)),0)`, s.d. = `r round(sd(sample_sums(ps7_genus)),0)`

#### Statistical methods for mixed-effects models
* Linear mixed-effects models fit with family ID as a random effect   
* Effect of country determined by comparing: (i) model with country/age vs (ii) model with age alone    
* Pairwise comparisons between countries (M vs I, U vs I, U vs M) corrected by FDR to control family error rate
* OPV/IPV arms from India combined in primary comparison
* Diversity in maternal samples compared by country via ANOVA with post-hoc Tukey test   





Taxon abundance profile {.hidden}
=====================================
Column {.tabset .tabset-fade}
-------------------------------------

### Genus-level

#### India - infant
```{r, fig.width=4, fig.height=4}
# for each sample subset, create prevalence vs abundance plot using taxon_abundance_function 
p = taxon_abundance_function(rps8_genus, genus=TRUE, input_country="India", input_colour=India_col)
ggMarginal(p, type = "density", fill=India_col, alpha=0.5, colour=NA, size=8)
```

N: `r nrow(subset(sample_data(rps8_genus), country=="India" & sample_type!="MS1"))`

#### Malawi - infant
```{r, fig.width=4, fig.height=4}
p = taxon_abundance_function(rps8_genus, genus=TRUE, input_country="Malawi", input_colour=Malawi_col)
ggMarginal(p, type = "density", fill=Malawi_col, alpha=0.5, colour=NA, size=8)
```

N: `r nrow(subset(sample_data(rps8_genus), country=="Malawi" & sample_type!="MS1"))`

#### UK - infant
```{r, fig.width=4, fig.height=4}
p = taxon_abundance_function(rps8_genus, genus=TRUE, input_country="UK", input_colour=UK_col)
ggMarginal(p, type = "density", fill=UK_col, alpha=0.5, colour=NA, size=8)
```

N: `r nrow(subset(sample_data(rps8_genus), country=="UK" & sample_type!="MS1"))`

#### India - maternal
```{r, fig.width=4, fig.height=4}
p = taxon_abundance_function(rps8_genus, genus=TRUE, input_country="India", input_colour=India_col, MS1=TRUE)
ggMarginal(p, type = "density", fill=India_col, alpha=0.5, colour=NA, size=8)
```

#### Malawi - maternal
```{r, fig.width=4, fig.height=4}
p = taxon_abundance_function(rps8_genus, genus=TRUE, input_country="Malawi", input_colour=Malawi_col, MS1=TRUE)
ggMarginal(p, type = "density", fill=Malawi_col, alpha=0.5, colour=NA, size=8)
```

#### UK - maternal
```{r, fig.width=4, fig.height=4}
p = taxon_abundance_function(rps8_genus, genus=TRUE, input_country="UK", input_colour=UK_col, MS1=TRUE)
ggMarginal(p, type = "density", fill=UK_col, alpha=0.5, colour=NA, size=8)
```

### RSV-level

#### India - infant
```{r, fig.width=4, fig.height=4}
p = taxon_abundance_function(rps8, genus=FALSE, input_country="India", input_colour=India_col)
ggMarginal(p, type = "density", fill=India_col, alpha=0.5, colour=NA, size=8)
```

#### Malawi - infant
```{r, fig.width=4, fig.height=4}
p = taxon_abundance_function(rps8, genus=FALSE, input_country="Malawi", input_colour=Malawi_col)
ggMarginal(p, type = "density", fill=Malawi_col, alpha=0.5, colour=NA, size=8)
```

#### UK - infant
```{r, fig.width=4, fig.height=4}
p = taxon_abundance_function(rps8, genus=FALSE, input_country="UK", input_colour=UK_col)
ggMarginal(p, type = "density", fill=UK_col, alpha=0.5, colour=NA, size=8)
```

#### India - maternal
```{r, fig.width=4, fig.height=4}
p = taxon_abundance_function(rps8, genus=FALSE, input_country="India", input_colour=India_col, MS1=TRUE)
ggMarginal(p, type = "density", fill=India_col, alpha=0.5, colour=NA, size=8)
```

#### Malawi - maternal
```{r, fig.width=4, fig.height=4}
p = taxon_abundance_function(rps8, genus=FALSE, input_country="Malawi", input_colour=Malawi_col, MS1=TRUE)
ggMarginal(p, type = "density", fill=Malawi_col, alpha=0.5, colour=NA, size=8)
```

#### UK - maternal
```{r, fig.width=4, fig.height=4}
p = taxon_abundance_function(rps8, genus=FALSE, input_country="UK", input_colour=UK_col, MS1=TRUE)
ggMarginal(p, type = "density", fill=UK_col, alpha=0.5, colour=NA, size=8)
```





Alpha diversity {.hidden}
=====================================
Column {.tabset .tabset-fade}
-------------------------------------

### Genus-level

#### India vs Malawi vs UK - Shannon
```{r, fig.width=7, fig.height=3}
# create dataframe of infant samples
t = data.frame(sample_data(rps8_genus))

#t_BS3 = subset(t, sample_type == "BS3" & country!="UK")
#write.csv(t_BS3, "t_BS3.csv")

# updated coding for week
t$week_fig[t$week_fig=="week 1"] = "week of life 1"
t$week_fig[t$week_fig=="week 4"] = "week of life 4"
t$week_fig[t$week_fig=="week 6/8"] = "week of life 6"
t$week_fig[t$week_fig=="week 10/12"] = "week of life 10"
t$week_fig = factor(t$week_fig, levels = c("week of life 1","week of life 4","week of life 6","week of life 10"))

# updated coding for country
t$country[t$country=="India"] = "IND"
t$country[t$country=="Malawi"] = "MLW"
t$country[t$country=="UK"] = "UK"

# select infant samples
t1 = subset(t, sample_type!="MS1")

# add-ons for longitudinal plots
gg_add = function(plot) { plot +
    geom_jitter(size = 1, alpha = 0.2, width=0.25) +
    geom_boxplot(alpha = 0.5, outlier.shape=NA) + 
    scale_color_manual("country", values = c(India_col, Malawi_col, UK_col)) +
    xlab("") + theme(legend.position = "none") + facet_grid(.~week_fig) + theme(strip.background = element_blank()) +
    theme(axis.text = element_text(size = 13), strip.text = element_text(size = 13), axis.title = element_text(size = 13))
}

# create longitudinal plots for alpha diversity by country
#p1 = gg_add(ggplot(t1, aes(x = factor(week_fig), y = Shannon_rsv, group=family_ID, colour=factor(country))) + ylab("Shannon (variant)") + ylim(0,5)) 
gg_add(ggplot(t1, aes(x = factor(country), y = Shannon, colour=factor(country))) + ylab("Shannon"))
#grid.arrange(p1, p2, ncol=2)
```

#### India vs Malawi vs UK - richness
```{r, fig.width=7, fig.height=3}
# create longitudinal plots for alpha diversity by country
#p1 = gg_add(ggplot(t1, aes(x = factor(week_fig), y = Shannon_rsv, group=family_ID, colour=factor(country))) + ylab("Shannon (variant)") + ylim(0,5)) 
gg_add(ggplot(t1, aes(x = factor(country), y = Observed, colour=factor(country))) + ylab("richness")) + ylim(0,30)
#grid.arrange(p1, p2, ncol=2)
```

#### India vs Malawi vs UK with cross-sectional p values - Shannon
```{r, fig.width=8, fig.height=3}
p1 = ggplot(subset(t1, week_fig=="week of life 1"), aes(factor(country), y=Shannon, colour=factor(country))) + geom_jitter(size = 1.5, alpha = 0.2, width=0.3) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) + ylab("Shannon") + xlab("") + theme(legend.position = "none", strip.background = element_blank()) + 
    scale_colour_manual(values = c("IND" = India_col, "MLW" = Malawi_col, "UK" = UK_col)) + facet_grid(.~week_fig) +  ylim(0,3.5) +
    geom_signif(comparisons = list(c("IND", "UK")), annotations="ns", y_position = 3.25, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") + 
    geom_signif(comparisons = list(c("MLW", "UK")), annotations="***", y_position = 3, tip_length=0, vjust=0.5, size = 0.3, textsize=6, color=Malawi_col) +
    geom_signif(comparisons = list(c("MLW", "IND")), annotations="***", y_position = 2.75, tip_length=0, vjust=0.5, size = 0.3, textsize=6, color=Malawi_col) +
    theme(axis.text = element_text(size = 13), strip.text = element_text(size = 13), axis.title = element_text(size = 13))

p2 = ggplot(subset(t1, week_fig=="week of life 4"), aes(factor(country), y=Shannon, colour=factor(country))) + geom_jitter(size = 1.5, alpha = 0.2, width=0.3) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) + ylab("") + xlab("") + theme(legend.position = "none", strip.background = element_blank()) + 
    scale_colour_manual(values = c("IND" = India_col, "MLW" = Malawi_col, "UK" = UK_col)) + facet_grid(.~week_fig) +  ylim(0,3.5) +
    geom_signif(comparisons = list(c("IND", "UK")), annotations="ns", y_position = 3.25, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") + 
    geom_signif(comparisons = list(c("MLW", "UK")), annotations="ns", y_position = 3, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") +
    geom_signif(comparisons = list(c("MLW", "IND")), annotations="***", y_position = 2.75, tip_length=0, vjust=0.5, size = 0.3, textsize=6, color=Malawi_col) +
    theme(axis.text = element_text(size = 13), axis.text.y = element_text(color="white"), strip.text = element_text(size = 13), axis.title = element_text(size = 13))

p3 = ggplot(subset(t1, week_fig=="week of life 6"), aes(factor(country), y=Shannon, colour=factor(country))) + geom_jitter(size = 1.5, alpha = 0.2, width=0.3) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) + ylab("") + xlab("") + theme(legend.position = "none", strip.background = element_blank()) + 
    scale_colour_manual(values = c("IND" = India_col, "MLW" = Malawi_col, "UK" = UK_col)) + facet_grid(.~week_fig) +  ylim(0,3.5) +
    geom_signif(comparisons = list(c("IND", "UK")), annotations="ns", y_position = 3.25, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") + 
    geom_signif(comparisons = list(c("MLW", "UK")), annotations="ns", y_position = 3, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") +
    geom_signif(comparisons = list(c("MLW", "IND")), annotations="ns", y_position = 2.75, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") +
    theme(axis.text = element_text(size = 13), axis.text.y = element_text(color="white"), strip.text = element_text(size = 13), axis.title = element_text(size = 13))

p4 = ggplot(subset(t1, week_fig=="week of life 10"), aes(factor(country), y=Shannon, colour=factor(country))) + geom_jitter(size = 1.5, alpha = 0.2, width=0.3) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) + ylab("") + xlab("") + theme(legend.position = "none", strip.background = element_blank()) + 
    scale_colour_manual(values = c("IND" = India_col, "MLW" = Malawi_col, "UK" = UK_col)) + facet_grid(.~week_fig) + ylim(0,3.5) +
    geom_signif(comparisons = list(c("IND", "UK")), annotations="ns", y_position = 3.25, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") + 
    geom_signif(comparisons = list(c("MLW", "UK")), annotations="ns", y_position = 3, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") +
    geom_signif(comparisons = list(c("MLW", "IND")), annotations="ns", y_position = 2.75, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") +
    theme(axis.text = element_text(size = 13), axis.text.y = element_text(color="white"), strip.text = element_text(size = 13), axis.title = element_text(size = 13))

grid.arrange(p1, p2, p3, p4, ncol=4, padding= unit(0, "line"))
```

#### India vs Malawi vs UK with cross-sectional p values - richness
```{r, fig.width=8, fig.height=3}
p1 = ggplot(subset(t1, week_fig=="week of life 1"), aes(factor(country), y=Observed, colour=factor(country))) + geom_jitter(size = 1.5, alpha = 0.2, width=0.3) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) + ylab("richness") + xlab("") + theme(legend.position = "none", strip.background = element_blank()) + 
    scale_colour_manual(values = c("IND" = India_col, "MLW" = Malawi_col, "UK" = UK_col)) + facet_grid(.~week_fig) +  ylim(0,30) +
    geom_signif(comparisons = list(c("IND", "UK")), annotations="ns", y_position = 30, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") + 
    geom_signif(comparisons = list(c("MLW", "UK")), annotations="***", y_position = 27.5, tip_length=0, vjust=0.5, size = 0.3, textsize=6, color=Malawi_col) +
    geom_signif(comparisons = list(c("MLW", "IND")), annotations="***", y_position = 25, tip_length=0, vjust=0.5, size = 0.3, textsize=6, color=Malawi_col) +
    theme(axis.text = element_text(size = 13), strip.text = element_text(size = 13), axis.title = element_text(size = 13))

p2 = ggplot(subset(t1, week_fig=="week of life 4"), aes(factor(country), y=Observed, colour=factor(country))) + geom_jitter(size = 1.5, alpha = 0.2, width=0.3) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) + ylab("") + xlab("") + theme(legend.position = "none", strip.background = element_blank()) + 
    scale_colour_manual(values = c("IND" = India_col, "MLW" = Malawi_col, "UK" = UK_col)) + facet_grid(.~week_fig) +  ylim(0,30) +
    geom_signif(comparisons = list(c("IND", "UK")), annotations="ns", y_position = 30, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") + 
    geom_signif(comparisons = list(c("MLW", "UK")), annotations="*", y_position = 27.5, tip_length=0, vjust=0.5, size = 0.3, textsize=6, color=Malawi_col) +
    geom_signif(comparisons = list(c("MLW", "IND")), annotations="***", y_position = 25, tip_length=0, vjust=0.5, size = 0.3, textsize=6, color=Malawi_col) +
    theme(axis.text = element_text(size = 13), axis.text.y = element_text(color="white"), strip.text = element_text(size = 13), axis.title = element_text(size = 13))

p3 = ggplot(subset(t1, week_fig=="week of life 6"), aes(factor(country), y=Observed, colour=factor(country))) + geom_jitter(size = 1.5, alpha = 0.2, width=0.3) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) + ylab("") + xlab("") + theme(legend.position = "none", strip.background = element_blank()) + 
    scale_colour_manual(values = c("IND" = India_col, "MLW" = Malawi_col, "UK" = UK_col)) + facet_grid(.~week_fig) +  ylim(0,30) +
    geom_signif(comparisons = list(c("IND", "UK")), annotations="ns", y_position = 30, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") + 
    geom_signif(comparisons = list(c("MLW", "UK")), annotations="ns", y_position = 27.5, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") +
    geom_signif(comparisons = list(c("MLW", "IND")), annotations="***", y_position = 25, tip_length=0, vjust=0.5, size = 0.3, textsize=6, color=Malawi_col) +
    theme(axis.text = element_text(size = 13), axis.text.y = element_text(color="white"), strip.text = element_text(size = 13), axis.title = element_text(size = 13))

p4 = ggplot(subset(t1, week_fig=="week of life 10"), aes(factor(country), y=Observed, colour=factor(country))) + geom_jitter(size = 1.5, alpha = 0.2, width=0.3) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) + ylab("") + xlab("") + theme(legend.position = "none", strip.background = element_blank()) + 
    scale_colour_manual(values = c("IND" = India_col, "MLW" = Malawi_col, "UK" = UK_col)) + facet_grid(.~week_fig) + ylim(0,30) +
    geom_signif(comparisons = list(c("IND", "UK")), annotations="ns", y_position = 30, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") + 
    geom_signif(comparisons = list(c("MLW", "UK")), annotations="ns", y_position = 27.5, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") +
    geom_signif(comparisons = list(c("MLW", "IND")), annotations="***", y_position = 25, tip_length=0, vjust=0.5, size = 0.3, textsize=6, color=Malawi_col) +
    theme(axis.text = element_text(size = 13), axis.text.y = element_text(color="white"), strip.text = element_text(size = 13), axis.title = element_text(size = 13))

grid.arrange(p1, p2, p3, p4, ncol=4, padding= unit(0, "line"))
```

#### N by timepoint
```{r}
table(t1$sample_type, t1$country)
```

#### p values, longitudinal mixed-effects models (FDR-adjusted, 4 decimal places)
```{r}
# calculate lme models for pairwise comparisons using lme_alpha
p_shan = c(lme_alpha(subset(t1, country!="UK"), "Shannon"), lme_alpha(subset(t1, country!="MLW"), "Shannon"), lme_alpha(subset(t1, country!="IND"), "Shannon"))
p_obs = c(lme_alpha(subset(t1, country!="UK"), "log(Observed)"), lme_alpha(subset(t1, country!="MLW"), "log(Observed)"), lme_alpha(subset(t1, country!="IND"), "log(Observed)"))

results = data.frame(round(p.adjust(p_shan, method="fdr"),4), round(p.adjust(p_obs, method="fdr"),4))
names(results) = c("Shannon", "Richness")
rownames(results) = c("MLW-IND", "UK-IND", "UK-MLW")
results
```

#### p values, India (OPV) vs India (IPV)
```{r}
# calculate lme models for pairwise comparisons using lme_alpha
p_shan = lme_alpha(subset(t1, country=="IND"), "Shannon", comparison="country_arm")
p_obs = lme_alpha(subset(t1, country=="IND"), "log(Observed)", comparison="country_arm")

results = data.frame(round(p_shan,4), round(p_obs,4))
names(results) = c("Shannon", "Richness")
rownames(results) = c("India (OPV)-India (IPV)")
results
```

#### p values, cross-sectional models 
##### week of life 1
```{r}
t1 = subset(t, sample_type=="BS1")
p_shan = TukeyHSD(x=aov(t1$Shannon ~ t1$country), "t1$country", conf.level=0.95)$`t1$country`[,4]
p_obs = TukeyHSD(x=aov(log(t1$Observed) ~ t1$country), "t1$country", conf.level=0.95)$`t1$country`[,4]
results = data.frame(round(p_shan, 4), round(p_obs, 4))
names(results) = c("Shannon", "Richness")
results
```

##### week of life 4
```{r}
t1 = subset(t, sample_type=="BS2")
p_shan = TukeyHSD(x=aov(t1$Shannon ~ t1$country), "t1$country", conf.level=0.95)$`t1$country`[,4]
p_obs = TukeyHSD(x=aov(log(t1$Observed) ~ t1$country), "t1$country", conf.level=0.95)$`t1$country`[,4]
results = data.frame(round(p_shan, 4), round(p_obs, 4))
names(results) = c("Shannon", "Richness")
results
```

##### week of life 6
```{r}
t1 = subset(t, sample_type=="BS3")
p_shan = TukeyHSD(x=aov(t1$Shannon ~ t1$country), "t1$country", conf.level=0.95)$`t1$country`[,4]
p_obs = TukeyHSD(x=aov(log(t1$Observed) ~ t1$country), "t1$country", conf.level=0.95)$`t1$country`[,4]
results = data.frame(round(p_shan, 4), round(p_obs, 4))
names(results) = c("Shannon", "Richness")
results
```

##### week of life 10
```{r}
t1 = subset(t, sample_type=="BS5")
p_shan = TukeyHSD(x=aov(t1$Shannon ~ t1$country), "t1$country", conf.level=0.95)$`t1$country`[,4]
p_obs = TukeyHSD(x=aov(log(t1$Observed) ~ t1$country), "t1$country", conf.level=0.95)$`t1$country`[,4]
results = data.frame(round(p_shan, 4), round(p_obs, 4))
names(results) = c("Shannon", "Richness")
results
```


#### Maternal samples - Shannon
```{r, fig.width=3, fig.height=4}
t1 = subset(t, sample_type=="MS1")

ggplot(t1, aes(factor(country), y=Shannon, colour=factor(country))) + geom_jitter(size = 1.5, alpha = 0.2, width=0.3) +
    geom_boxplot(alpha = 0.5, outlier.shape=NA) + ylab("Shannon") + xlab("") + theme(legend.position = "none", strip.background = element_blank()) + #ylim(0,90) +
    scale_colour_manual(values = c("IND" = India_col, "MLW" = Malawi_col, "UK" = UK_col)) + 
    geom_signif(comparisons = list(c("IND", "UK")), annotations="***", y_position = 4.5, tip_length=0, vjust=0.5, size = 0.3, textsize=6, color=India_col) + 
    geom_signif(comparisons = list(c("MLW", "UK")), annotations="**", y_position = 4.25, tip_length=0, vjust=0.5, size = 0.3, textsize=6, color=Malawi_col) +
    geom_signif(comparisons = list(c("MLW", "IND")), annotations="ns", y_position = 4, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") +
   # scale_y_continuous(limits=c(0,225), breaks=c(0,50,100,150,200)) +
    theme(axis.text = element_text(size = 13), strip.text = element_text(size = 13), axis.title = element_text(size = 13))
```

#### Maternal samples - richness
```{r, fig.width=3, fig.height=4}
t1 = subset(t, sample_type=="MS1")

ggplot(t1, aes(factor(country), y=Observed, colour=factor(country))) + geom_jitter(size = 1.5, alpha = 0.2, width=0.3) +
    geom_boxplot(alpha = 0.5, outlier.shape=NA) + ylab("richness") + xlab("") + theme(legend.position = "none", strip.background = element_blank()) + ylim(0,90) +
    scale_colour_manual(values = c("IND" = India_col, "MLW" = Malawi_col, "UK" = UK_col)) + 
    geom_signif(comparisons = list(c("IND", "UK")), annotations="***", y_position = 87.5, tip_length=0, vjust=0.5, size = 0.3, textsize=6, color=India_col) + 
    geom_signif(comparisons = list(c("MLW", "UK")), annotations="***", y_position = 81.25, tip_length=0, vjust=0.5, size = 0.3, textsize=6, color=Malawi_col) +
    geom_signif(comparisons = list(c("MLW", "IND")), annotations="ns", y_position = 75, tip_length=0, vjust=0, size = 0.3, textsize=4, color="grey") +
   # scale_y_continuous(limits=c(0,225), breaks=c(0,50,100,150,200)) +
    theme(axis.text = element_text(size = 13), strip.text = element_text(size = 13), axis.title = element_text(size = 13))
```

#### p values, India vs Malawi vs UK
```{r}
t1 = subset(t, sample_type=="MS1")
p_shan = TukeyHSD(x=aov(t1$Shannon ~ t1$country), "t1$country", conf.level=0.95)$`t1$country`[,4]
p_obs = TukeyHSD(x=aov(t1$Observed ~ t1$country), "t1$country", conf.level=0.95)$`t1$country`[,4]
results = data.frame(round(p_shan, 4), round(p_obs, 4))
names(results) = c("Shannon", "Richness")
results
```

#### p values, India (OPV) vs India (IPV)
```{r}
t1 = subset(t, sample_type=="MS1" & country=="IND")
p_shan = TukeyHSD(x=aov(t1$Shannon ~ t1$country_arm), "t1$country_arm", conf.level=0.95)$`t1$country_arm`[,4]
p_obs = TukeyHSD(x=aov(t1$Observed ~ t1$country_arm), "t1$country_arm", conf.level=0.95)$`t1$country_arm`[,4]
results = data.frame(round(p_shan, 4), round(p_obs, 4))
names(results) = c("Shannon", "Richness")
results
```





Longitudinal abundance plots {.hidden}
===================================== 

#### Infants - India vs UK vs Malawi
```{r, fig.width=8, fig.height=4}
g = data.frame(t(otu_table(rps8_genus))) 

# add metadata for sample type and country
if (all(rownames(g)==sample_names(rps8_genus))) { 
  g$sample_type =  as.factor(sample_data(rps8_genus)$sample_type)
  g$country_arm =  as.factor(sample_data(rps8_genus)$country_arm)
} else { print("error: rownames do not match")}

# collate data
g_means_full = data.frame(
  BS1_I = genus_mean_cal(g, "BS1", "India"),
  BS2_I = genus_mean_cal(g, "BS2", "India"),
  BS3_I = genus_mean_cal(g, "BS3", "India"),
  BS5_I = genus_mean_cal(g, "BS5", "India"),
  BS1_M = genus_mean_cal(g, "BS1", "Malawi"),
  BS2_M = genus_mean_cal(g, "BS2", "Malawi"),
  BS3_M = genus_mean_cal(g, "BS3", "Malawi"),
  BS5_M = genus_mean_cal(g, "BS5", "Malawi"),
  BS1_U = genus_mean_cal(g, "BS1", "UK"),
  BS2_U = genus_mean_cal(g, "BS2", "UK"),
  BS3_U = genus_mean_cal(g, "BS3", "UK"),
  BS5_U = genus_mean_cal(g, "BS5", "UK"),
  BS1_II = genus_mean_cal(g, "BS1", "India (IPV)"),
  BS2_II = genus_mean_cal(g, "BS2", "India (IPV)"),
  BS3_II = genus_mean_cal(g, "BS3", "India (IPV)"),
  BS5_II = genus_mean_cal(g, "BS5", "India (IPV)"),
  BS1_IO = genus_mean_cal(g, "BS1", "India (OPV)"),
  BS2_IO = genus_mean_cal(g, "BS2", "India (OPV)"),
  BS3_IO = genus_mean_cal(g, "BS3", "India (OPV)"),
  BS5_IO = genus_mean_cal(g, "BS5", "India (OPV)"),
  MS1_I = genus_mean_cal(g, "MS1", "India"),
  MS1_M = genus_mean_cal(g, "MS1", "Malawi"),
  MS1_U = genus_mean_cal(g, "MS1", "UK")
)

# create subset for Indian, Malawian and UK infant samples (columns 1:12)
g_means = g_means_full[,1:12]

# calculate means, pick genera present with mean abundance of 5% during at least one timepoint
g_means$max = apply(g_means, 1, max)
g_means$include = g_means$max>=0.05
g_means_reduced = subset(g_means, include==TRUE)[,1:(ncol(g_means)-2)]
taxa_list = rownames(g_means_reduced)
g_means_reduced["Other",] = 1-colSums(g_means_reduced)
g_means_reduced = data.frame(t(g_means_reduced))

# collate data for each taxon in taxa_list
g_collated = rbind(data.frame(Genus = rep("Other",12), week = rep(c("1", "4", "6", "10"),3), 
                              country = c(rep("IND",4), rep("MLW",4), rep("UK",4)), abundance = g_means_reduced[,"Other"]),
                   data.frame(Genus = rep("Parabacteroides",12), week = rep(c("1", "4", "6", "10"),3), 
                              country = c(rep("IND",4), rep("MLW",4), rep("UK",4)), abundance = g_means_reduced[,"Parabacteroides"]),
                   data.frame(Genus = rep("Prevotella 9",12), week = rep(c("1", "4", "6", "10"),3), 
                              country = c(rep("IND",4), rep("MLW",4), rep("UK",4)), abundance = g_means_reduced[,"Prevotella_9"]),
                   data.frame(Genus = rep("Klebsiella",12), week = rep(c("1", "4", "6", "10"),3), 
                              country = c(rep("IND",4), rep("MLW",4), rep("UK",4)), abundance = g_means_reduced[,"Klebsiella"]),
                   data.frame(Genus = rep("Lactobacillus",12), week = rep(c("1", "4", "6", "10"),3), 
                              country = c(rep("IND",4), rep("MLW",4), rep("UK",4)), abundance = g_means_reduced[,"Lactobacillus"]),
                   data.frame(Genus = rep("Veillonella",12), week = rep(c("1", "4", "6", "10"),3), 
                              country = c(rep("IND",4), rep("MLW",4), rep("UK",4)), abundance = g_means_reduced[,"Veillonella"]),
                   data.frame(Genus = rep("Bacteroides",12), week = rep(c("1", "4", "6", "10"),3), 
                              country = c(rep("IND",4), rep("MLW",4), rep("UK",4)), abundance = g_means_reduced[,"Bacteroides"]),
                   data.frame(Genus = rep("Staphylococcus",12), week = rep(c("1", "4", "6", "10"),3), 
                              country = c(rep("IND",4), rep("MLW",4), rep("UK",4)), abundance = g_means_reduced[,"Staphylococcus"]),
                   data.frame(Genus = rep("Streptococcus",12), week = rep(c("1", "4", "6", "10"),3), 
                              country = c(rep("IND",4), rep("MLW",4), rep("UK",4)), abundance = g_means_reduced[,"Streptococcus"]),
                   data.frame(Genus = rep("Escherichia/Shigella",12), week = rep(c("1", "4", "6", "10"),3), 
                              country = c(rep("IND",4), rep("MLW",4), rep("UK",4)), abundance = g_means_reduced[,"Escherichia.Shigella"]),
                   data.frame(Genus = rep("Bifidobacterium",12), week = rep(c("1", "4", "6", "10"),3), 
                              country = c(rep("IND",4), rep("MLW",4), rep("UK",4)), abundance = g_means_reduced[,"Bifidobacterium"]))

# recode week
g_collated$week_factor <- factor(g_collated$week, levels = c("1", "4", "6", "10"))
g_collated$week <- as.numeric(g_collated$week_factor)
names(g_collated)[1] = "genus"

# create plot
ggplot(g_collated, aes(x=week, y=abundance, fill=genus)) +
  geom_area() + facet_grid(.~country) + 
  scale_x_continuous(breaks = 1:4, labels = c("1", "4", "6", "10")) +
  scale_fill_manual(values = c("light grey",as.vector(GetColors(11,"smooth rainbow")))) +
  ylab("mean relative abundance") + xlab("week of life") +
  scale_y_continuous(breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  theme(strip.background = element_blank()) + 
  theme(axis.text = element_text(size=13), strip.text = element_text(size=13), axis.title=element_text(size=13),
        legend.title = element_text(size=13), legend.text = element_text(size=13))
```

#### N by timepoint
```{r}
table(sample_data(rps8_genus)$sample_type, sample_data(rps8_genus)$country)
```

#### Infants - India (OPV) vs India (IPV)
```{r, fig.width=6, fig.height=4}
# create subset for Indian OPV and IPV arms (columns 13:20)
g_means = g_means_full[,13:20]
g_means_reduced = g_means[taxa_list,]
g_means_reduced["Other",] = 1-colSums(g_means_reduced)
g_means_reduced = data.frame(t(g_means_reduced))

# collate data for each taxon in taxa_list
g_collated = rbind(data.frame(Genus = rep("Other",8), week = rep(c("1", "4", "6", "10"),2), 
                              country = c(rep("IND (IPV)",4), rep("IND (OPV)",4)), abundance = g_means_reduced[,"Other"]),
                   data.frame(Genus = rep("Parabacteroides",8), week = rep(c("1", "4", "6", "10"),2), 
                              country = c(rep("IND (IPV)",4), rep("IND (OPV)",4)), abundance = g_means_reduced[,"Parabacteroides"]),
                   data.frame(Genus = rep("Prevotella 9",8), week = rep(c("1", "4", "6", "10"),2), 
                              country = c(rep("IND (IPV)",4), rep("IND (OPV)",4)), abundance = g_means_reduced[,"Prevotella_9"]),
                   data.frame(Genus = rep("Klebsiella",8), week = rep(c("1", "4", "6", "10"),2), 
                              country = c(rep("IND (IPV)",4), rep("IND (OPV)",4)), abundance = g_means_reduced[,"Klebsiella"]),
                   data.frame(Genus = rep("Lactobacillus",8), week = rep(c("1", "4", "6", "10"),2), 
                              country = c(rep("IND (IPV)",4), rep("IND (OPV)",4)), abundance = g_means_reduced[,"Lactobacillus"]),
                   data.frame(Genus = rep("Veillonella",8), week = rep(c("1", "4", "6", "10"),2), 
                              country = c(rep("IND (IPV)",4), rep("IND (OPV)",4)), abundance = g_means_reduced[,"Veillonella"]),
                   data.frame(Genus = rep("Bacteroides",8), week = rep(c("1", "4", "6", "10"),2), 
                              country = c(rep("IND (IPV)",4), rep("IND (OPV)",4)), abundance = g_means_reduced[,"Bacteroides"]),
                   data.frame(Genus = rep("Staphylococcus",8), week = rep(c("1", "4", "6", "10"),2), 
                              country = c(rep("IND (IPV)",4), rep("IND (OPV)",4)), abundance = g_means_reduced[,"Staphylococcus"]),
                   data.frame(Genus = rep("Streptococcus",8), week = rep(c("1", "4", "6", "10"),2), 
                              country = c(rep("IND (IPV)",4), rep("IND (OPV)",4)), abundance = g_means_reduced[,"Streptococcus"]),
                   data.frame(Genus = rep("Escherichia/Shigella",8), week = rep(c("1", "4", "6", "10"),2), 
                              country = c(rep("IND (IPV)",4), rep("IND (OPV)",4)), abundance = g_means_reduced[,"Escherichia.Shigella"]),
                   data.frame(Genus = rep("Bifidobacterium",8), week = rep(c("1", "4", "6", "10"),2), 
                              country = c(rep("IND (IPV)",4), rep("IND (OPV)",4)), abundance = g_means_reduced[,"Bifidobacterium"]))

# recode week
g_collated$week_factor <- factor(g_collated$week, levels = c("1", "4", "6", "10"))
g_collated$week <- as.numeric(g_collated$week_factor)
names(g_collated)[1] = "genus"

# create plot
ggplot(g_collated, aes(x=week, y=abundance, fill=genus)) +
  geom_area() + facet_grid(.~country) + 
  scale_x_continuous(breaks = 1:4, labels = c("1", "4", "6", "10")) +
  scale_y_continuous(breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  scale_fill_manual(values = c("light grey",as.vector(GetColors(11,"smooth rainbow")))) +
  ylab("mean relative abundance") + 
  theme(strip.background = element_blank()) +
  theme(axis.text = element_text(size=13), strip.text = element_text(size=13), axis.title=element_text(size=13),
        legend.title = element_text(size=13), legend.text = element_text(size=13))
```

#### Maternal samples (with genera corresponding to infant plots included)
```{r, fig.width=5, fig.height=4}
# create subset for maternal samples (columns 21:23)
g_means = g_means_full[,21:23]
g_means_reduced = g_means[taxa_list,]
g_means_reduced["Other",] = 1-colSums(g_means_reduced)
g_means_reduced = data.frame(t(g_means_reduced))

# collate data for each taxon in taxa_list
g_collated = rbind(data.frame(Genus = rep("Other",3), week = rep(c("mother"),3), country = c("IND", "MLW", "UK"), abundance = g_means_reduced[,"Other"]),
                   data.frame(Genus = rep("Parabacteroides",3), week = rep(c("mother"),3), country = c("IND", "MLW", "UK"), abundance = g_means_reduced[,"Parabacteroides"]),
                   data.frame(Genus = rep("Prevotella 9",3), week = rep(c("mother"),3), country = c("IND", "MLW", "UK"), abundance = g_means_reduced[,"Prevotella_9"]),
                   data.frame(Genus = rep("Klebsiella",3), week = rep(c("mother"),3), country = c("IND", "MLW", "UK"), abundance = g_means_reduced[,"Klebsiella"]),
                   data.frame(Genus = rep("Lactobacillus",3), week = rep(c("mother"),3), country = c("IND", "MLW", "UK"), abundance = g_means_reduced[,"Lactobacillus"]),
                   data.frame(Genus = rep("Veillonella",3), week = rep(c("mother"),3), country = c("IND", "MLW", "UK"), abundance = g_means_reduced[,"Veillonella"]),
                   data.frame(Genus = rep("Bacteroides",3), week = rep(c("mother"),3), country = c("IND", "MLW", "UK"), abundance = g_means_reduced[,"Bacteroides"]),
                   data.frame(Genus = rep("Staphylococcus",3), week = rep(c("mother"),3), country = c("IND", "MLW", "UK"), abundance = g_means_reduced[,"Staphylococcus"]),
                   data.frame(Genus = rep("Streptococcus",3), week = rep(c("mother"),3), country = c("IND", "MLW", "UK"), abundance = g_means_reduced[,"Streptococcus"]),
                   data.frame(Genus = rep("Escherichia/Shigella",3), week = rep(c("mother"),3), country = c("IND", "MLW", "UK"), abundance = g_means_reduced[,"Escherichia.Shigella"]),
                   data.frame(Genus = rep("Bifidobacterium",3), week = rep(c("mother"),3), country = c("IND", "MLW", "UK"), abundance = g_means_reduced[,"Bifidobacterium"]))
names(g_collated)[1] = "genus"

# create plot
ggplot(g_collated, aes(x=country, y=abundance, fill=genus)) + geom_bar(stat = "identity") + xlab("") +
  scale_fill_manual(values = c("light grey",as.vector(GetColors(11,"smooth rainbow")))) + 
  ylab("mean relative abundance") + 
  scale_y_continuous(breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  theme(axis.text = element_text(size=13), strip.text = element_text(size=13), axis.title=element_text(size=13),
        legend.title = element_text(size=13), legend.text = element_text(size=13))
```





Beta diversity {.hidden}
=====================================
Column {.tabset .tabset-fade}
-------------------------------------

### PERMANOVA - unweighted

##### Inter-individual differences among all infant samples
```{r}
# inter-individual
ps_t = subset_samples(rps8_genus, sample_type=="BS1" | sample_type=="BS2" | sample_type=="BS3" | sample_type=="BS5")
t1 = data.frame(sample_data(ps_t))
otu_table(ps_t)[otu_table(ps_t)<0.001] = 0
beta_dist = phyloseq::distance(ps_t, method = "bray", binary=TRUE) 
adon_BS = adonis(beta_dist ~ family_ID, data = t1, permutations = adonis_iter)$aov.tab
adon_BS
```

##### Differences by country stratified by age
```{r}
# BS1
ps_t = subset_samples(rps8_genus, sample_type=="BS1")
t1 = data.frame(sample_data(ps_t))
otu_table(ps_t)[otu_table(ps_t)<0.001] = 0
beta_dist = phyloseq::distance(ps_t, method = "bray", binary=TRUE) 
adon_BS1 = adonis(beta_dist ~ country, data = t1, permutations = adonis_iter)$aov.tab

# BS2
ps_t = subset_samples(rps8_genus, sample_type=="BS2")
t2 = data.frame(sample_data(ps_t))
otu_table(ps_t)[otu_table(ps_t)<0.001] = 0
beta_dist = phyloseq::distance(ps_t, method = "bray", binary=TRUE) 
adon_BS2 = adonis(beta_dist ~ country, data = t2, permutations = adonis_iter)$aov.tab

# BS3
ps_t = subset_samples(rps8_genus, sample_type=="BS3")
t3 = data.frame(sample_data(ps_t))
otu_table(ps_t)[otu_table(ps_t)<0.001] = 0
beta_dist = phyloseq::distance(ps_t, method = "bray", binary=TRUE) 
adon_BS3 = adonis(beta_dist ~ country, data = t3, permutations = adonis_iter)$aov.tab

# BS5
ps_t = subset_samples(rps8_genus, sample_type=="BS5")
t5 = data.frame(sample_data(ps_t))
otu_table(ps_t)[otu_table(ps_t)<0.001] = 0
beta_dist = phyloseq::distance(ps_t, method = "bray", binary=TRUE) 
adon_BS5 = adonis(beta_dist ~ country, data = t5, permutations = adonis_iter)$aov.tab

# MS1
ps_t = subset_samples(rps8_genus, sample_type=="MS1")
tm = data.frame(sample_data(ps_t))
otu_table(ps_t)[otu_table(ps_t)<0.001] = 0
beta_dist = phyloseq::distance(ps_t, method = "bray", binary=TRUE) 
adon_MS1 = adonis(beta_dist ~ country, data = tm, permutations = adonis_iter)$aov.tab


collated = data.frame(
  week = c("week of life 1", "week of life 4", "week of life 6", "week of life 10", "mother"),
  R2 = c(adon_BS1$R2[1], adon_BS2$R2[1], adon_BS3$R2[1], adon_BS5$R2[1], adon_MS1$R2[1]),
  p =  c(adon_BS1$`Pr(>F)`[1], adon_BS2$`Pr(>F)`[1], adon_BS3$`Pr(>F)`[1], adon_BS5$`Pr(>F)`[1], adon_MS1$`Pr(>F)`[1]),
  n = c(nrow(t1), nrow(t2), nrow(t3), nrow(t5), nrow(tm))
)
collated
```

#### Summary plot
```{r, fig.height=3, fig.width=4}
collated$week = factor(collated$week, levels =  c("week of life 1", "week of life 4", "week of life 6", "week of life 10", "mother"))
ggplot(collated, aes(y = R2*100, x=week, fill = week )) + geom_bar(stat = "identity") +
  scale_x_discrete(limits = rev(levels(collated$week))) + xlab("") + ylab(bquote(R^2~"(%)")) +
  scale_y_continuous(limits=c(0,25), breaks=c(0,10,20)) + 
  coord_flip() + theme(legend.position = "none") + scale_fill_manual(values = rep("#0072B2",5)) +
  theme(plot.title = element_text(size=15), axis.text = element_text(size=15), axis.title = element_text(size=15), strip.text = element_text(size=15)) + 
  theme(strip.background = element_blank())
```

### PERMANOVA - weighted

##### Inter-individual differences among all infant samples
```{r}
# inter-individual
ps_t = subset_samples(rps8_genus, sample_type=="BS1" | sample_type=="BS2" | sample_type=="BS3" | sample_type=="BS5")
t1 = data.frame(sample_data(ps_t))
beta_dist = phyloseq::distance(ps_t, method = "bray") 
adon_BS = adonis(beta_dist ~ family_ID, data = t1, permutations = adonis_iter)$aov.tab
adon_BS
```

##### Differences by country stratified by age
```{r}
# BS1
ps_t = subset_samples(rps8_genus, sample_type=="BS1")
t1 = data.frame(sample_data(ps_t))
beta_dist = phyloseq::distance(ps_t, method = "bray") 
adon_BS1 = adonis(beta_dist ~ country, data = t1, permutations = adonis_iter)$aov.tab

# BS2
ps_t = subset_samples(rps8_genus, sample_type=="BS2")
t2 = data.frame(sample_data(ps_t))
beta_dist = phyloseq::distance(ps_t, method = "bray") 
adon_BS2 = adonis(beta_dist ~ country, data = t2, permutations = adonis_iter)$aov.tab

# BS3
ps_t = subset_samples(rps8_genus, sample_type=="BS3")
t3 = data.frame(sample_data(ps_t))
beta_dist = phyloseq::distance(ps_t, method = "bray") 
adon_BS3 = adonis(beta_dist ~ country, data = t3, permutations = adonis_iter)$aov.tab

# BS5
ps_t = subset_samples(rps8_genus, sample_type=="BS5")
t5 = data.frame(sample_data(ps_t))
beta_dist = phyloseq::distance(ps_t, method = "bray") 
adon_BS5 = adonis(beta_dist ~ country, data = t5, permutations = adonis_iter)$aov.tab

# MS1
ps_t = subset_samples(rps8_genus, sample_type=="MS1")
tm = data.frame(sample_data(ps_t))
beta_dist = phyloseq::distance(ps_t, method = "bray") 
adon_MS1 = adonis(beta_dist ~ country, data = tm, permutations = adonis_iter)$aov.tab


collated = data.frame(
  week = c("week of life 1", "week of life 4", "week of life 6", "week of life 10", "mother"),
  R2 = c(adon_BS1$R2[1], adon_BS2$R2[1], adon_BS3$R2[1], adon_BS5$R2[1], adon_MS1$R2[1]),
  p =  c(adon_BS1$`Pr(>F)`[1], adon_BS2$`Pr(>F)`[1], adon_BS3$`Pr(>F)`[1], adon_BS5$`Pr(>F)`[1], adon_MS1$`Pr(>F)`[1]),
  n = c(nrow(t1), nrow(t2), nrow(t3), nrow(t5), nrow(tm))
)
collated
```

#### Summary plot
```{r, fig.height=3, fig.width=4}
collated$week = factor(collated$week, levels =  c("week of life 1", "week of life 4", "week of life 6", "week of life 10", "mother"))
ggplot(collated, aes(y = R2*100, x=week, fill = week )) + geom_bar(stat = "identity") +
  scale_x_discrete(limits = rev(levels(collated$week))) + xlab("") + ylab(bquote(R^2~"(%)")) +
  scale_y_continuous(limits=c(0,25), breaks=c(0,10,20)) + 
  coord_flip() + theme(legend.position = "none") + scale_fill_manual(values = rep("#0072B2",5)) +
  theme(plot.title = element_text(size=15), axis.text = element_text(size=15), axis.title = element_text(size=15), strip.text = element_text(size=15)) + 
  theme(strip.background = element_blank())
```

#### Turnover, paired vs random (weighted)
For week 1, distance from maternal sample plotted. For subsequent samples, distance from prior infant sample plotted.
```{r, fig.width=9, fig.height=7}
# create dataframe, and recode week
t1 = subset(t, sample_type!="MS1")
within = data.frame(sample_ID_full = t1$sample_ID_full, country = t1$country, week_fig = t1$week_fig, turnover=t1$turnover_w, within = rep("paired", nrow(t1)))
between = data.frame(sample_ID_full = t1$sample_ID_full, country = t1$country, week_fig = t1$week_fig, turnover=t1$turnover_randompair_w, within = rep("random", nrow(t1)))
full = rbind(within, between)
full$week_fig = factor(full$week_fig, levels = c("week of life 1","week of life 4","week of life 6","week of life 10"))

# create plot
ggplot(full, aes(factor(within), turnover, colour=factor(within))) + geom_jitter(size = 1, alpha = 0.2, width=0.2) +
  geom_boxplot(alpha = 0.5) + facet_grid(country~week_fig) + 
  ylab("turnover (weighted)") + xlab("") + scale_color_manual(values = c("#01665e", "#d8b365")) + 
  stat_compare_means(label.y = 1.1, size=4) + 
  scale_y_continuous(limits = c(0,1.2), breaks=c(0,0.4,0.8,1.2)) +
  theme_bw() + theme(legend.position = "none", strip.background = element_blank()) + 
  theme(axis.text = element_text(size = 13), axis.title = element_text(size = 13), plot.title = element_text(size = 13), strip.text = element_text(size = 13))
```

#### Turnover, paired vs random (unweighted)
```{r, fig.width=9, fig.height=7}
# create dataframe, and recode week
t1 = subset(t, sample_type!="MS1")
within = data.frame(sample_ID_full = t1$sample_ID_full, country = t1$country, week_fig = t1$week_fig, turnover=t1$turnover_u, within = rep("paired", nrow(t1)))
between = data.frame(sample_ID_full = t1$sample_ID_full, country = t1$country, week_fig = t1$week_fig, turnover=t1$turnover_randompair_u, within = rep("random", nrow(t1)))
full = rbind(within, between)
full$week_fig = factor(full$week_fig, levels = c("week of life 1","week of life 4","week of life 6","week of life 10"))

# create plot
ggplot(full, aes(factor(within), turnover, colour=factor(within))) + geom_jitter(size = 1, alpha = 0.2, width=0.2) +
  geom_boxplot(alpha = 0.5) + facet_grid(country~week_fig) + 
  ylab("turnover (weighted)") + xlab("") + scale_color_manual(values = c("#01665e", "#d8b365")) + 
  stat_compare_means(label.y = 1.1, size=4) + 
  scale_y_continuous(limits = c(0,1.2), breaks=c(0,0.4,0.8,1.2)) +
  theme_bw() + theme(legend.position = "none", strip.background = element_blank()) + 
  theme(axis.text = element_text(size = 13), axis.title = element_text(size = 13), plot.title = element_text(size = 13), strip.text = element_text(size = 13))
```
Longitudinal samples from a given infant more closely clustered than expected by chance in each of the three cohorts.





Differential taxa {.hidden}
=====================================
Column {.tabset .tabset-fade}
-------------------------------------

### Genus-level
```{r}
set.seed(12345)

### select input data and create specify output comparisons
input_m = data.frame(sample_data(rps8_genus))
comparison_list = data.frame(country1 = c("India", "India", "Malawi", "India (OPV)"),
                             country2 = c("Malawi", "UK", "UK", "India (IPV)"))
subset_list = c("BS1", "BS2", "BS3", "BS5", "MS1")

# create dataframe for collated accuracy results
collated_rf = data.frame(subset = rep(NA,5*4))
collated_rf$subset = collated_rf$country1 = collated_rf$country2 = collated_rf$comparison = collated_rf$crossval_median = collated_rf$crossval_q1 = collated_rf$crossval_q3 = NA

if (run_rf_full) {
  # run RF in loop across 5 sample types and 4 country-country comparisons
  for (s in 1:5) {
    for (c in 1:4) {
      subset = subset_list[s]
      country1 = comparison_list$country1[c]
      country2 = comparison_list$country2[c]
      if (country1=="India (OPV)") {
        samplelist = subset(input_m, (country_arm==country1 | country_arm==country2) & sample_type==subset)$sample_ID_full
      } else {
        samplelist = subset(input_m, (country==country1 | country==country2) & sample_type==subset)$sample_ID_full
      }
      input_rps = prune_samples(samplelist, rps8_genus)
      input_ps = prune_samples(samplelist, ps7_genus)
      if (country1=="India") { 
        sample_data(input_rps)$country_arm = sample_data(input_rps)$country 
        sample_data(input_ps)$country_arm = sample_data(input_ps)$country 
        }
      rf = run_rf(input_rps, input_ps)

      collated_rf$subset[4*(s-1)+c] = subset
      collated_rf$country1[4*(s-1)+c] = as.character(country1)
      collated_rf$country2[4*(s-1)+c] = as.character(country2)
      collated_rf$comparison[4*(s-1)+c] = paste0(country1,"-",country2)
      collated_rf$crossval_median[4*(s-1)+c] = rf$gini.cv$crossval_median[1]
      collated_rf$crossval_q1[4*(s-1)+c] = rf$gini.cv$crossval_quartile_1[1]
      collated_rf$crossval_q3[4*(s-1)+c] = rf$gini.cv$crossval_quartile_3[1]
    }
  }
  write.csv(collated_rf,"output_module3/crossval_collated_country_genus.csv")
}
collated_rf_full = read.csv("output_module3/crossval_collated_country_genus.csv", row.names=1)
```

#### Cross-validation accuracy, India vs Malawi vs UK
```{r, fig.width=10, fig.height=3.5}
collated_rf_full$week = revalue(as.factor(collated_rf_full$subset), c("BS1" = "week of life 1", "BS2" = "week of life 4", "BS3" = "week of life 6", "BS5" = "week of life 10", "MS1" = "mother"))
collated_rf = subset(collated_rf_full, country2!="India (IPV)")

# recode comparison
collated_rf$comparison = as.character(collated_rf$comparison)
collated_rf$comparison[collated_rf$comparison=="India-Malawi"] = "IND-MLW"
collated_rf$comparison[collated_rf$comparison=="India-UK"] = "IND-UK"
collated_rf$comparison[collated_rf$comparison=="Malawi-UK"] = "MLW-UK"

ggplot(collated_rf, aes(y = crossval_median, x = comparison, colour=factor(comparison))) + 
  geom_point(size = 5, alpha=0.8) +
  geom_errorbar(aes(ymin=crossval_q1, ymax=crossval_q3), width=0) + 
  facet_grid(.~week) + geom_hline(yintercept = 0.5, linetype="dotted") +
  scale_y_continuous(limits=c(0.4, 1), oob=rescale_none, breaks=c(0.5,0.75,1)) + #coord_flip() +
  scale_colour_manual(values = as.vector(GetColors(8,"sunset", reverse=T))) +
  theme(legend.position = "") + xlab("") + ylab("CV accuracy") + 
  theme(strip.background = element_blank()) +
  theme(axis.text.x = element_text(size = 15, angle = 90, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 18), axis.title = element_text(size = 18), plot.title = element_text(size = 13), 
        strip.text = element_text(size = 18))
```

#### Pre-vaccination comparisons
```{r, fig.width=12, fig.height=17}
IM_BS3 = read.csv("output_module3/RF_output_importance/crossval_gini_India_Malawi_BS3_genus.csv", row.names=1)
IU_BS3 = read.csv("output_module3/RF_output_importance/crossval_gini_India_UK_BS3_genus.csv", row.names=1)
MU_BS3 = read.csv("output_module3/RF_output_importance/crossval_gini_Malawi_UK_BS3_genus.csv", row.names=1)

gini = plot_grid(gini_plot(IM_BS3, India_col, Malawi_col, species=F) + ggtitle("week of life 6") + ylab(" ") + scale_y_continuous(breaks = c(0,1,2), limits = c(0,2)),
                 gini_plot(IU_BS3, India_col, UK_col, species=F) + ggtitle(" ") + ylab(" ") + scale_y_continuous(breaks = c(0,1,2), limits = c(0,2)),
                 gini_plot(MU_BS3, Malawi_col, UK_col,species=F) + ggtitle(" ") + scale_y_continuous(breaks = c(0,1,2), limits = c(0,2)) + 
                   theme(axis.title = element_text(size=22)),
                 ncol=1, align = "v")
prev = plot_grid(rf_prev_plot(IM_BS3, label1="India", colour1 = India_col, 
                              label2 = "Malawi", colour2=Malawi_col, species=F) + ggtitle(" ") + ylab(" "),
                 rf_prev_plot(IU_BS3, label1="India", colour1 = India_col, 
                              label2 = "UK", colour2=UK_col, species=F) + ggtitle(" ") + ylab(" "),
                 rf_prev_plot(MU_BS3, label1="Malawi", colour1 = Malawi_col, 
                              label2 = "UK", colour2=UK_col, species=F) + ggtitle(" ") + theme(axis.title = element_text(size=22)), ncol=1, align = "v")
grid.arrange(gini, prev, ncol=2, widths=c(2,1.4))
```

#### Differential abundance plots - significant by Fisher's test or Aldex2 (FDR p <0.05)
```{r, fig.width=9, fig.height=6}
plot_grid(
  summary_abund_plot("India","IND", "Malawi", "MLW", India_col, Malawi_col, "genus", threshold=0.05, yshift=3, ymax=100) + xlab(" ") + ylab("number of genera") +
    scale_y_continuous(limits = c(0,100), breaks = c(0,25,50,75)),
  summary_abund_plot("India", "IND", "UK", "UK", India_col, UK_col, "genus", threshold=0.05, yshift=3, ymax=100) + ylab(" ") + 
    scale_y_continuous(limits = c(0,100), breaks = c(0,25,50,75)),
  summary_abund_plot("Malawi", "MLW", "UK", "UK", Malawi_col, UK_col, "genus", threshold=0.05, yshift=3, ymax=100) + ylab(" ") + xlab(" ") +
    scale_y_continuous(limits = c(0,100), breaks = c(0,25,50,75)), 
  ncol=3)
```

Longitudinal abundance comparisons {.hidden}
=====================================
Column {.tabset .tabset-fade}
-------------------------------------

### Genus-level
```{r}
# select subsets of data
t = data.frame(as(tax_table(ps7_genus), "matrix"))
t_rsv = data.frame(as(tax_table(ps7_l_t_pr_d_dd_25k), "matrix"))

# append sequence of most common rsv to each genus
t$sequence = NA
for (i in 1:nrow(t)) {
  genus = as.character(t$Genus[i]) # pick out genus for each row in turn
  t_rsv_sub = subset(t_rsv, Genus==genus) # find matching RSVs
  t$sequence[i] = as.character(t_rsv_sub$sequence[1]) # pick out top sequence (previously ordered by overall abundance)
}
#all(rownames(t) == taxa_names(ps6_genus))
t1 = tax_table(t)
rownames(t1) = rownames(t)
colnames(t1) = names(t)
tax_table(ps7_genus) = t1

# pick out subset of taxa present in at least 25% of infant samples in at least one country
ps7_I = subset_samples(ps7_genus, sample_type!="MS1" & country=="India") %>%
  prune_taxa(rowSums(otu_table(.)>0)/nsamples(.)>=0.20, .)
ps7_U = subset_samples(ps7_genus, sample_type!="MS1" & country_arm=="UK") %>%
  prune_taxa(rowSums(otu_table(.)>0)/nsamples(.)>=0.20, .)
ps7_M = subset_samples(ps7_genus, sample_type!="MS1" & country_arm=="Malawi") %>%
  prune_taxa(rowSums(otu_table(.)>0)/nsamples(.)>=0.20, .)
taxa_list = unique(c(taxa_names(ps7_I), taxa_names(ps7_U), taxa_names(ps7_M)))
ps7_t = prune_taxa(taxa_list, ps7_genus)

# determine mean abundances
rps7_t = transform_sample_counts(ps7_t, function(x) {x/sum(x)})
rps7_t = subset_samples(rps7_t, sample_type!="MS1")
means = data.frame(genus = tax_table(rps7_t)[,"Genus"], means = rowMeans(otu_table(rps7_t)))
names(means)[1] = "genus"
```

#### India vs UK
```{r}
if (run_nb_full) {
  # pick out subset of samples and taxa
  ps7_t1 = subset_samples(ps7_t, sample_type!="MS1" & (country=="India" | country=="UK"))
  ps7_t1 = prune_taxa(c(taxa_names(ps7_I), taxa_names(ps7_U)), ps7_t1)

  # create feature table and pick out metadata variables
  features = data.frame(t(as(otu_table(ps7_t1), "matrix")))
  if (all(colnames(features) == rownames(tax_table(ps7_t1)))) { 
    colnames(features) = data.frame(as(tax_table(ps7_t1), "matrix"))$Genus 
    }
  meta = data.frame(sample_data(ps7_t1))
  N = meta$count_ps7     
  week = meta$week
  country = meta$country
  subject = meta$family_ID

  # ZINB models (all taxa)
  f = mms(y = features, fixed = ~ country + week + offset(log(N)), 
          random = ~ 1 | subject, min.p = 0, method = "zinb", verbose=TRUE)
  res = data.frame(get.fixed(f, part="dist", vr.name="countryUK", sort.p=F))
  res$method = "zinb"
  
  # negative binomials without zero inflation (if prevalence >95%)
  f_nb <- mms(y = features, fixed = ~ country + week + offset(log(N)), 
              random = ~ 1 | subject, min.p = 0.95, method = "nb", verbose=TRUE)
  res_nb = data.frame(get.fixed(f_nb, part="dist", vr.name="countryUK", sort.p=F))
  res_nb$method = "nb"
  
  # replace ZINB with negative binomial model outputs for relevant taxa
  for (i in 1:nrow(res_nb)){
      if (rownames(res_nb)[i] %in% rownames(res)) { 
        res[rownames(res_nb)[i],] = res_nb[i,] 
        } else { res = rbind(res, res_nb[i,]) }
    }
  
  # collate results
  res$genus = rownames(res)
  res$padj = p.adjust(res$pvalue, method="BH")
  res$signif = "ns"
  res$signif[res$padj<0.05 & res$Estimate<0]  = "IND"
  res$signif[res$padj<0.05 & res$Estimate>0]  = "UK"
  res$pscale = -log10(res$padj)
  res$n_tested = ncol(features)
  res$failed_convergence = ncol(features)-nrow(res)
  
  # sample counts
  meta1 = subset(meta, country=="India")
  meta2 = subset(meta, country=="UK")
  res$nsamples_country1 = nrow(meta1)
  res$nsamples_country2 = nrow(meta2)
  res$ninfants_country1 = length(unique(meta1$family_ID))
  res$ninfants_country2 = length(unique(meta2$family_ID))

  res$comparison = "IND-UK"
  res_I_U = res
} else {
  res = read.csv("output_module3/longitudinal_zinb_country.csv", header=TRUE, stringsAsFactors=FALSE)[,-1]
  res_I_U = subset(res, comparison == "IND-UK")
}
```

Number of genera tested: `r res_I_U$n_tested[1]`    
Number of discriminant genera (FDR p < 0.05): `r sum(res_I_U$padj<0.05)` (India = `r sum(res_I_U$signif=="IND")`; UK = `r sum(res_I_U$signif=="UK")`)
Failed to converge: `r res_I_U$failed_convergence[1]`

#### India vs Malawi
```{r}
if (run_nb_full) {
  # pick out subset of samples and taxa
  ps7_t1 = subset_samples(ps7_t, sample_type!="MS1" & (country=="India" | country=="Malawi"))
  ps7_t1 = prune_taxa(c(taxa_names(ps7_I), taxa_names(ps7_M)), ps7_t1)

  # create feature table and pick out metadata variables
  features = data.frame(t(as(otu_table(ps7_t1), "matrix")))
  if (all(colnames(features) == rownames(tax_table(ps7_t1)))) { 
    colnames(features) = data.frame(as(tax_table(ps7_t1), "matrix"))$Genus 
    }
  meta = data.frame(sample_data(ps7_t1))
  N = meta$count_ps7     
  week = meta$week
  country = meta$country
  subject = meta$family_ID

  # ZINB models (all taxa)
  f = mms(y = features, fixed = ~ country + week + offset(log(N)), 
          random = ~ 1 | subject, min.p = 0, method = "zinb", verbose=TRUE)
  res = data.frame(get.fixed(f, part="dist", vr.name="countryMalawi", sort.p=F))
  res$method = "zinb"
  
  # negative binomials without zero inflation (if prevalence >95%)
  f_nb <- mms(y = features, fixed = ~ country + week + offset(log(N)), 
              random = ~ 1 | subject, min.p = 0.95, method = "nb", verbose=TRUE)
  res_nb = data.frame(get.fixed(f_nb, part="dist", vr.name="countryMalawi", sort.p=F))
  res_nb$method = "nb"
  
  # replace ZINB with negative binomial model outputs for relevant taxa
  for (i in 1:nrow(res_nb)){
      if (rownames(res_nb)[i] %in% rownames(res)) { 
        res[rownames(res_nb)[i],] = res_nb[i,] 
        } else { res = rbind(res, res_nb[i,]) }
  }
  
  # collate results
  res$genus = rownames(res)
  res$padj = p.adjust(res$pvalue, method="BH")
  res$signif = "ns"
  res$signif[res$padj<0.05 & res$Estimate<0]  = "IND"
  res$signif[res$padj<0.05 & res$Estimate>0]  = "MLW"
  res$pscale = -log10(res$padj)
  res$n_tested = ncol(features)
  res$failed_convergence = ncol(features)-nrow(res)
  
  # sample counts
  meta1 = subset(meta, country=="India")
  meta2 = subset(meta, country=="Malawi")
  res$nsamples_country1 = nrow(meta1)
  res$nsamples_country2 = nrow(meta2)
  res$ninfants_country1 = length(unique(meta1$family_ID))
  res$ninfants_country2 = length(unique(meta2$family_ID))

  res$comparison = "IND-MLW"
  res_I_M = res
} else {
  res_I_M = subset(res, comparison == "IND-MLW")
}
```
Number of genera tested: `r res_I_M$n_tested[1]`    
Number of discriminant genera (FDR p < 0.05): `r sum(res_I_M$padj<0.05)` (India = `r sum(res_I_M$signif=="IND")`; Malawi = `r sum(res_I_M$signif=="MLW")`)
Failed to converge: `r res_I_M$failed_convergence[1]`

#### Malawi vs UK
```{r}
if (run_nb_full) {
  # pick out subset of samples and taxa
  ps7_t1 = subset_samples(ps7_t, sample_type!="MS1" & (country=="Malawi" | country=="UK"))
  ps7_t1 = prune_taxa(c(taxa_names(ps7_M), taxa_names(ps7_U)), ps7_t1)

  # create feature table and pick out metadata variables
  features = data.frame(t(as(otu_table(ps7_t1), "matrix")))
  if (all(colnames(features) == rownames(tax_table(ps7_t1)))) { 
    colnames(features) = data.frame(as(tax_table(ps7_t1), "matrix"))$Genus 
    }
  meta = data.frame(sample_data(ps7_t1))
  N = meta$count_ps7     
  week = meta$week
  country = meta$country
  subject = meta$family_ID

  # ZINB models (all taxa)
  f = mms(y = features, fixed = ~ country + week + offset(log(N)), 
          random = ~ 1 | subject, min.p = 0, method = "zinb", verbose=TRUE)
  res = data.frame(get.fixed(f, part="dist", vr.name="countryUK", sort.p=F))
  res$method = "zinb"
  
  # negative binomials without zero inflation (if prevalence >95%)
  f_nb <- mms(y = features, fixed = ~ country + week + offset(log(N)), 
              random = ~ 1 | subject, min.p = 0.95, method = "nb", verbose=TRUE)
  res_nb = data.frame(get.fixed(f_nb, part="dist", vr.name="countryUK", sort.p=F))
  res_nb$method = "nb"
  
  # replace ZINB with negative binomial model outputs for relevant taxa
  for (i in 1:nrow(res_nb)){
      if (rownames(res_nb)[i] %in% rownames(res)) { 
        res[rownames(res_nb)[i],] = res_nb[i,] 
        } else { res = rbind(res, res_nb[i,]) }
    }
  
  # collate results
  res$genus = rownames(res)
  res$padj = p.adjust(res$pvalue, method="BH")
  res$signif = "ns"
  res$signif[res$padj<0.05 & res$Estimate<0]  = "MLW"
  res$signif[res$padj<0.05 & res$Estimate>0]  = "UK"
  res$pscale = -log10(res$padj)
  res$n_tested = ncol(features)
  res$failed_convergence = ncol(features)-nrow(res)
  
  # sample counts
  meta1 = subset(meta, country=="Malawi")
  meta2 = subset(meta, country=="UK")
  res$nsamples_country1 = nrow(meta1)
  res$nsamples_country2 = nrow(meta2)
  res$ninfants_country1 = length(unique(meta1$family_ID))
  res$ninfants_country2 = length(unique(meta2$family_ID))

  res$comparison = "MLW-UK"
  res_M_U = res
  res = rbind(res_I_U, res_I_M, res_M_U)
  write.csv(res, "output_module3/longitudinal_zinb_country.csv")
} else {
  res_M_U = subset(res, comparison == "MLW-UK")
}
```
Number of genera tested: `r res_M_U$n_tested[1]`    
Number of discriminant genera (FDR p < 0.05): `r sum(res_M_U$padj<0.05)` (Malawi = `r sum(res_M_U$signif=="MLW")`; UK = `r sum(res_M_U$signif=="UK")`)
Failed to converge: `r res_M_U$failed_convergence[1]`

#### Construct tree
```{r, fig.height=10, fig.width=10, warnings=FALSE}
# append mean abundances 
res = merge(means, res, by = "genus")

# pick out sequences
seqs = as.character(data.frame(as(tax_table(ps7_t), "matrix"))$sequence)
names(seqs) = seqs 

# create alignment
alignment = AlignSeqs(DNAStringSet(seqs), anchor=NA,verbose=FALSE) # align
phangAlign <- phyDat(as(alignment, "matrix"), type="DNA")
dm <- dist.ml(phangAlign)

# create neighbour-joining tree
treeNJ = NJ(dm) # note that tip order != sequence order

# re-assign tip labels by genus name
if (all(treeNJ$tip.label == as.character(data.frame(as(tax_table(ps7_t), "matrix"))$sequence))) {
  treeNJ$tip.label =  as.character(data.frame(as(tax_table(ps7_t), "matrix"))$Genus)
}

# create tree metadata
tree_m <- data.frame(taxa = treeNJ$tip.label)
tree_m$phylum = NA
tax_ref = data.frame(as(tax_table(ps7_t), "matrix"))
for (i in 1:nrow(tree_m)) { 
  tree_m$phylum[i] = as.character(tax_ref$Phylum[tax_ref$Genus==tree_m$taxa[i]]) 
}

#treeNJ$tip.label[treeNJ$tip.label=="Escherichia/Shigella"] = "Escherichia"
treeNJ$tip.label = gsub("_", " ", treeNJ$tip.label) 
treeNJ$tip.label = gsub("unassigned ", "", treeNJ$tip.label) 
tree_m$taxa = gsub("_", " ", tree_m$taxa) 
tree_m$taxa = gsub("unassigned ", "", tree_m$taxa) 
res$genus = gsub("_", " ", res$genus) 
res$genus = gsub("unassigned ", "", res$genus) 


# draw tree
t = ggtree(treeNJ) 
t1 = t %<+% tree_m + geom_tippoint(aes(color=phylum), shape = 15, alpha=0.9, size=2) + 
  scale_colour_manual(values = c("#332288","#56B4E9","#999999","#D55E00","#999933")) +
  geom_tiplab(size=4, align=TRUE, linesize=.2, offset = 0.2, aes(color=phylum)) + 
  theme(legend.position = "left", legend.title = element_text(size=13), legend.text = element_text(size=13)) + xlim(0, 2)

# pick out tip order
d = fortify(treeNJ)
d = subset(d, isTip)
order = data.frame(genus = with(d, label[order(y, decreasing=T)]))
order$branch_order = factor(1:nrow(order))
res1 = merge(res, order, by = "genus")

# specify upper and lower limits for estimate and p values
res1$Estimate[res1$Estimate>=10] = 10
res1$Estimate[res1$Estimate<=(-10)] = -10
res1$pscale[res1$pscale>=50] = 50
res1$Estimate[res1$comparison=="MLW-UK"] = (-1)*res1$Estimate[res1$comparison=="MLW-UK"]
res1$abundance = "abund."

# generate plot of abundance circles
g0 = ggplot(res1, aes(x = rep(0, nrow(res)), 
                      y = reorder(branch_order, dplyr::desc(branch_order)), size = asin(sqrt(means)))) +
  geom_point(alpha=0.9, color = "#d95f0e", shape = 1) + theme_minimal() + 
  theme(axis.title.y = element_blank(), axis.text.y=element_blank(), 
        axis.text.x=element_blank(), legend.position = "", plot.title = element_text(size=11, hjust = 0.5)) + 
  theme(plot.margin = unit(c(0.1,0,0.35,0), "cm"), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("") + scale_size(limits = c(0, 0.8)) + facet_grid(.~abundance) +
  theme(strip.text = element_text(size=13), legend.text = element_text(size=13), legend.title= element_text(size=13), axis.text= element_text(size=13))

# generate plot of enriched/depleted taxa by country
g1 = ggplot(res1, aes(x = Estimate, y = reorder(branch_order, dplyr::desc(branch_order)), 
                      colour = signif, size = pscale)) + geom_point(alpha=0.9) + theme_minimal() + 
  theme(axis.title.y = element_blank(), axis.text.y=element_blank()) + 
  scale_color_manual(values = c(ns = "grey", IND = India_col, MLW = Malawi_col, UK = UK_col)) +
  xlim(-12,12) + xlab(" ") + 
  scale_x_continuous(limits=c(-10,10), breaks=c(-10,0,10)) +
  theme(legend.position = "right", plot.title = element_text(size=11, hjust = 0.5)) + 
  facet_grid(.~comparison) + theme(plot.margin = unit(c(0.1,0,-0.2,0), "cm")) +
  theme(panel.spacing = unit(1, "lines")) + labs(colour="enriched country", size = "-log10(p)") +
  theme(strip.text = element_text(size=13), legend.text = element_text(size=13), legend.title= element_text(size=13),
        axis.text= element_text(size=13), axis.title= element_text(size=13))

grid.arrange(t1, g0, g1, ncol=3, widths = c(1.8,0.3,1.8))
```

#### Longitudinal abundance plots for major genera
```{r, fig.width=10, fig.height=8, message=FALSE, warning=FALSE}
# select genus abundance data and append genus names
g = data.frame(t(otu_table(rps7_genus))) 
g$sample_ID_full = rownames(g)
t = data.frame(sample_data(rps7_genus))

# append metadata
g = merge(g, t, by = "sample_ID_full")
g$Other = 1 - (g$Bifidobacterium + g$Escherichia.Shigella + g$Streptococcus + g$Staphylococcus + g$Bacteroides + g$Veillonella + g$Lactobacillus + g$Klebsiella + g$Prevotella_9 + g$Parabacteroides)

g$country[g$country=="India"] = "IND"
g$country[g$country=="Malawi"] = "MLW"

# create genus plots
p1 = genus_plot(g, "Bifidobacterium") + ylab("relative abundance")
p2 = genus_plot(g, "Escherichia.Shigella")  + ggtitle("Escherichia/Shigella") + ylab("")
p3 = genus_plot(g, "Streptococcus") 
p4 = genus_plot(g, "Staphylococcus") 
p5 = genus_plot(g, "Bacteroides") + ylab("relative abundance")
p6 = genus_plot(g, "Veillonella")
p7 = genus_plot(g, "Lactobacillus")
p8 = genus_plot(g, "Klebsiella") + xlab("week of life")
p9 = genus_plot(g, "Prevotella_9") + ggtitle("Prevotella 9") + xlab("week of life")  + ylab("relative abundance")
p10 = genus_plot(g, "Parabacteroides") + xlab("week of life")
p11 = genus_plot(g, "Other") + ylab("") + xlab("week of life")
plot_grid(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11, ncol=4, align="vh")
```


Session info {.hidden}
===================================== 

```{r}
print(sessionInfo())
```
.
