---
title: "Metagenome vs demography, growth and birthmode"
author: "RoVI study"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  flexdashboard::flex_dashboard:
    self_contained: yes
    theme: flatly 
---

Sidebar {.sidebar} 
===================================== 
[Sample summary]    

[India]    

[Malawi]

[Session info]



Sample summary {.hidden}
=====================================

```{r, message=FALSE}
# variable to define whether to run analyses in full (TRUE) or not (FALSE)
run_beta_full = TRUE # set as TRUE to run beta diversity analyses in full

# load packages and functions
source("../packages_functions.R")

# load feature data - rarefied data for genus/species modules; otherwise full data
if (grepl("genus", getwd())) {
  load("final_ps_rare.RData")
  ps = ps_rare
} else if (grepl("species", getwd())) {
  load("final_ps_rare.RData")
  ps = ps_rare
} else if (grepl("pathway", getwd())) {
  load("final_ps_full.RData")
  ps = ps_full
} else {
  load("final_ps_full.RData")
  ps = ps_full
}

# transform to relative abundances
rps = transform_sample_counts(ps, function(x) {x/sum(x)})

# load metadata and recodeset factor levels
a = data.frame(sample_data(rps))
a$BS1_VP6_cat[!is.na(a$BS1_VP6_cat)] = factor(a$BS1_VP6_cat)[!is.na(a$BS1_VP6_cat)]
a$sex_baby = factor(a$sex_baby, levels=c("male", "female"))
a$crowding_wami = factor(a$crowding_wami, levels=c("up_to_4", "over_4"))
a$kitchen = factor(a$kitchen, levels=c("no", "yes"))
a$sanitation = factor(a$sanitation, levels=c("no", "yes"))
a$refrigerator = factor(a$refrigerator, levels=c("no", "yes"))
a$type_house = factor(a$type_house, levels=c("kutcha_mixed", "pucca"))
a$antibiotic_exposure = factor(a$antibiotic_exposure, levels=c("no", "yes"))
a$place_delivery = factor(a$place_delivery, levels=c("non_tertiary", "tertiary"))
a$mode_delivery = factor(a$mode_delivery, levels=c("vaginal", "caesarean"))
a$breastfed_child = factor(a$breastfed_child, levels=c("nonexclusive", "exclusive"))
a$pre_exposed = factor(a$pre_exposed)
levels(a$pre_exposed) <- c("no", "yes")

# recode 'unknown' as NA for water treatment status and HIV status
a$water_treat_wami[a$water_treat_wami=="unknown"] = NA
a$water_treat_wami = factor(a$water_treat_wami, levels=c("no", "yes"))
a$HIV_status[a$HIV_status=="unknown"] = NA
a$HIV_status = factor(a$HIV_status, levels=c("negative", "positive"))
sample_data(rps) = a
```

#### Input data: infant stools (BS3) 
* n samples = `r nsamples(rps)`    
* n taxa: genera = `r ntaxa(rps)`
* mean read count: `r round(mean(sample_sums(ps)),0)`, s.d. = `r round(sd(sample_sums(ps)),0)`    
<br><br>





India {.hidden}
=====================================

#### Beta diversity - unweighted Bray-Curtis distances
```{r, fig.width=8, fig.height=8}
system("mkdir exploratory_outputs")

### run beta diversity analyses
beta_vars = read.csv("../beta_variables_reduced.csv", stringsAsFactors = FALSE)
beta_vars = subset(beta_vars, IND==1)

if (run_beta_full) {
  rps_t = subset_samples(rps, country=="India")
  t = data.frame(sample_data(rps_t))
  ncomp = nrow(beta_vars)
  collated = data.frame(
    variable = beta_vars$BS3, 
    clean_variable = beta_vars$clean_name,
    group = beta_vars$group, 
    class = NA,
    beta_R2 = NA, 
    beta_p = NA, 
    n_samples = NA,
    alpha_test = NA,
    alpha_level1 = NA,
    alpha_level2 = NA,
    alpha_mean_sd1 = NA,
    alpha_mean_sd2 = NA,
    alpha_rho = NA,
    alpha_p = NA
  )
  for (i in 1:nrow(collated)) {
    if (!is.na(collated$variable[i])) {
      t$outcome = t[,as.character(collated$variable[i])]
      sample_list = as.character(t$sample_ID[!is.na(t$outcome)])
      rps_sub = prune_samples(sample_list, rps_t) %>% prune_taxa(taxa_sums(.) > 0, .)
      beta = data.frame(sample_data(rps_sub))
      beta$outcome = beta[,as.character(collated$variable[i])]
      collated$n_samples[i] = length(sample_list)
      if (class(t$outcome)=="character") {
        t$outcome = factor(t$outcome)
      }
      collated$class[i] = class(t$outcome)
    
      # unweighted adonis
      bray_dist_u = phyloseq::distance(rps_sub, method = "bray", binary = TRUE)
      adon = adonis(unname(bray_dist_u) ~ outcome, data = beta, permutations = 999)
      collated$beta_R2[i] = round(adon$aov.tab$R2[1],3)
      collated$beta_p[i] = round(adon$aov.tab$`Pr(>F)`[1],5)
      
      # Spearman's or wilcox test for alpha 
      if (class(t$outcome)=="integer" | class(t$outcome)=="numeric") {
        collated$alpha_test[i] = "spearman"
        collated$alpha_rho[i] = round(cor(t$outcome, t$Observed, method="spearman"),3)
        collated$alpha_p[i] = round(cor.test(t$outcome, t$Observed, method="spearman")$p.value,5)
      }
      if (class(t$outcome)=="factor") {
        collated$alpha_test[i] = "wilcoxon"
        collated$alpha_p[i] = round(wilcox.test(t$Observed ~ t$outcome)$p.value,5)
        alpha_summary = aggregate(Observed ~ outcome, t, function(x) c(mean = mean(x), sd = sd(x)))
        collated$alpha_level1[i] = as.character(alpha_summary[1,1])
        collated$alpha_level2[i] = as.character(alpha_summary[2,1])
        collated$alpha_mean_sd1[i] = paste0(round(alpha_summary[2][1,][1],1)," (",round(alpha_summary[2][1,][2],1),")")
        collated$alpha_mean_sd2[i] = paste0(round(alpha_summary[2][2,][1],1)," (",round(alpha_summary[2][2,][2],1),")")
      }
    }
  }
  collated$beta_fdr_p = round(p.adjust(collated$beta_p, method="fdr"),5)
  collated$alpha_fdr_p = round(p.adjust(collated$alpha_p, method="fdr"),5)
  write.csv(collated, "exploratory_outputs/India_collated.csv")
} else { collated = read.csv("exploratory_outputs/India_collated.csv") }

collated$clean_variable = factor(collated$clean_variable, levels = beta_vars$clean_name)
collated$clean_variable = revalue(collated$clean_variable, c("A1AT at 6w (μg/ml)"="α1AT at 6w (μg/ml)","A1AG at 6w (μg/ml)"="α1AG at 6w (μg/ml)"))
collated$group = factor(collated$group, levels = c("maternal", "household", "infant", "rotavirus"))
collated$group = as.character(collated$group)

### beta plot
ggplot(collated, aes(y = beta_R2*100, x=clean_variable, fill = group)) + 
  geom_bar(stat = "identity") + 
  scale_x_discrete(limits = rev(levels(collated$clean_variable))) + xlab("") + 
  ylab("R2 (%)") + scale_y_continuous(limits = c(0,8), breaks=c(0,4,8)) + 
  coord_flip() + theme(legend.position="none") +
 scale_fill_manual(values=c(household = "#CC79A7", infant = "#009E73", maternal = "#56B4E9", rotavirus = "#661100")) + ggtitle("PERMANOVA,\nunweighted Bray-Curtis") + 
  theme(axis.text.y = element_text(size=17), axis.text.x = element_text(size=18), axis.title = element_text(size=18), 
        legend.text = element_text(size=18), legend.title = element_text(size=18), plot.title = element_text(size=18)) 
```

##### Significant assocations - Bray-Curtis (FDR p <0.2)
```{r}
subset(collated, beta_fdr_p<0.2)[,c("clean_variable", "beta_R2", "beta_p", "beta_fdr_p", "n_samples")]
```

##### Significant assocations - Richness (FDR p <0.2)
```{r}
subset(collated, alpha_fdr_p<0.2)[,c("clean_variable", "alpha_test", "alpha_level1", "alpha_level2", "alpha_mean_sd1", "alpha_mean_sd2", "alpha_rho", "alpha_p", "alpha_fdr_p", "n_samples")]
```




Malawi {.hidden}
=====================================

#### Beta diversity - unweighted Bray-Curtis distances
```{r, fig.width=7, fig.height=5}
### run beta diversity analyses
beta_vars = read.csv("../beta_variables_reduced.csv", stringsAsFactors = FALSE)
beta_vars = subset(beta_vars, MLW==1)

if (run_beta_full) {
  rps_t = subset_samples(rps, country=="Malawi")
  t = data.frame(sample_data(rps_t))
  ncomp = nrow(beta_vars)
  collated = data.frame(
    variable = beta_vars$BS3, 
    clean_variable = beta_vars$clean_name,
    group = beta_vars$group, 
    class = NA,
    beta_R2 = NA, 
    beta_p = NA, 
    n_samples = NA,
    alpha_test = NA,
    alpha_level1 = NA,
    alpha_level2 = NA,
    alpha_mean_sd1 = NA,
    alpha_mean_sd2 = NA,
    alpha_rho = NA,
    alpha_p = NA
  )
  for (i in 1:nrow(collated)) {
    if (!is.na(collated$variable[i])) {
      t$outcome = t[,as.character(collated$variable[i])]
      sample_list = as.character(t$sample_ID[!is.na(t$outcome)])
      rps_sub = prune_samples(sample_list, rps_t) %>% prune_taxa(taxa_sums(.) > 0, .)
      beta = data.frame(sample_data(rps_sub))
      beta$outcome = beta[,as.character(collated$variable[i])]
      collated$n_samples[i] = length(sample_list)
      if (class(t$outcome)=="character") {
        t$outcome = factor(t$outcome)
      }
      collated$class[i] = class(t$outcome)
    
      # unweighted adonis
      bray_dist_u = phyloseq::distance(rps_sub, method = "bray", binary = TRUE)
      adon = adonis(unname(bray_dist_u) ~ outcome, data = beta, permutations = 999)
      collated$beta_R2[i] = round(adon$aov.tab$R2[1],3)
      collated$beta_p[i] = round(adon$aov.tab$`Pr(>F)`[1],5)
      
      # Spearman's or wilcox test for alpha 
      if (class(t$outcome)=="integer" | class(t$outcome)=="numeric") {
        collated$alpha_test[i] = "spearman"
        collated$alpha_rho[i] = round(cor(t$outcome, t$Observed, method="spearman"),3)
        collated$alpha_p[i] = round(cor.test(t$outcome, t$Observed, method="spearman")$p.value,5)
      }
      if (class(t$outcome)=="factor") {
        collated$alpha_test[i] = "wilcoxon"
        collated$alpha_p[i] = round(wilcox.test(t$Observed ~ t$outcome)$p.value,5)
        alpha_summary = aggregate(Observed ~ outcome, t, function(x) c(mean = mean(x), sd = sd(x)))
        collated$alpha_level1[i] = as.character(alpha_summary[1,1])
        collated$alpha_level2[i] = as.character(alpha_summary[2,1])
        collated$alpha_mean_sd1[i] = paste0(round(alpha_summary[2][1,][1],1)," (",round(alpha_summary[2][1,][2],1),")")
        collated$alpha_mean_sd2[i] = paste0(round(alpha_summary[2][2,][1],1)," (",round(alpha_summary[2][2,][2],1),")")
      }
    }
  }
  collated$beta_fdr_p = round(p.adjust(collated$beta_p, method="fdr"),5)
  collated$alpha_fdr_p = round(p.adjust(collated$alpha_p, method="fdr"),5)
  write.csv(collated, "exploratory_outputs/Malawi_collated.csv")
} else { collated = read.csv("exploratory_outputs/Malawi_collated.csv") }

collated$clean_variable = factor(collated$clean_variable, levels = beta_vars$clean_name)
collated$clean_variable = revalue(collated$clean_variable, c("A1AT at 6w (μg/ml)"="α1AT at 6w (μg/ml)","A1AG at 6w (μg/ml)"="α1AG at 6w (μg/ml)"))
collated$group = factor(collated$group, levels = c("maternal", "household", "infant", "rotavirus"))
collated$group = as.character(collated$group)

### beta plot
ggplot(collated, aes(y = beta_R2*100, x=clean_variable, fill = group)) + 
  geom_bar(stat = "identity") + 
  scale_x_discrete(limits = rev(levels(collated$clean_variable))) + xlab("") + 
  ylab("R2 (%)") + scale_y_continuous(limits = c(0,8), breaks=c(0,4,8)) + 
  coord_flip() + theme(legend.position="none") +
 scale_fill_manual(values=c(household = "#CC79A7", infant = "#009E73", maternal = "#56B4E9", rotavirus = "#661100")) + ggtitle("PERMANOVA,\nunweighted Bray-Curtis") + 
  theme(axis.text.y = element_text(size=17), axis.text.x = element_text(size=18), axis.title = element_text(size=18), 
        legend.text = element_text(size=18), legend.title = element_text(size=18), plot.title = element_text(size=18)) 
```

##### Significant assocations - Bray-Curtis (FDR p <0.2)
```{r}
subset(collated, beta_fdr_p<0.2)[,c("clean_variable", "beta_R2", "beta_p", "beta_fdr_p", "n_samples")]
```

##### Significant assocations - Richness (FDR p <0.2)
```{r}
subset(collated, alpha_fdr_p<0.2)[,c("clean_variable", "alpha_test", "alpha_level1", "alpha_level2", "alpha_mean_sd1", "alpha_mean_sd2", "alpha_rho", "alpha_p", "alpha_fdr_p", "n_samples")]
```





Session info {.hidden}
===================================== 

```{r}
print(sessionInfo())
```
.